```{r}
library(finalfit)
library(ggplot2)
library(tidyverse)
library(readxl)
library(lubridate)
library(corrplot)
library(naniar)
library(keras)
library(glmnet)
library(xgboost)
library(finalfit)
library(rpart)
library(MASS)
library(ranger)

WL.analysis <- read_csv("WL.analysis.csv")
head(WL.analysis)
```
Material

```{r}
ggplot(WL.analysis, aes(x = MATERIAL, y = condition_overall_score, fill = MATERIAL)) + geom_boxplot() + labs(
  title = "Relationship between Material and Condition",
  x = "Material",
  y = "Condition Score"
)
```

Age Analysis

```{r}
ggplot(WL.analysis, aes(x = INSTALLED, y = condition_overall_score, color = MATERIAL)) + geom_point() + labs(
  x = "Year of installation",
  y = "Condition Score",
  title = "Relationship between the installation year of the pipe and it's condition"
)



ggplot(WL.analysis, aes(x = INSTALLED, y = condition_overall_score, color = MATERIAL)) + geom_point() + geom_smooth(method = "lm", col = "blue") + labs(
  x = "Age of pipe",
  y = "Condition Score",
  title = "Relationship between the age of the pipe and it's condition"
)

```

```{r}
head(WL.analysis)
```


Time to first failure analysis

```{r}

WL.analysis.clean <- WL.analysis %>% filter(!is.na(INSTALLED), !is.na(`_concatenated_break_dates`))
result <- WL.analysis.clean %>%
  arrange(`_concatenated_break_dates`) %>% # 1. Arrange by earliest break date
  distinct(OBJECTID, MATERIAL, INSTALLED, NOM_DIA_MM, LENGTH, `_concatenated_break_dates`, condition_overall_score, condition_overall_score_label, 
           condition_leak_history_overall_score, condition_likelihood_of_failure_overall_score, condition_Confidence_Scenario, 
           constructability_overall_score, constructability_overall_score_label, constructability_environmental_scoring_dataset_s_, 
           criticality_overall_score, criticality_overall_score_label, criticality_key_customer_category, criticality_key_customer_score,
           criticality_capacity_score, criticality_customer_affected_score, criticality_pipe_ranking_score, criticality_other_critical_asset_score, criticality_other_critical_asset_dataset_s_, criticality_Confidence_Scenario, `Risk-score`,
           c.keep_all = TRUE) # 2. Remove duplicate pipes

head(result)
```


```{r}
# Calculate the difference between installation year and first break date
result$time_to_first_failure <- year(result$`_concatenated_break_dates`) - year(result$INSTALLED)

library(fitdistrplus)

ggplot(result, aes(x = time_to_first_failure)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "black", fill = "blue", position = "identity") +
  labs(title = "Distribution of Time to First Failure", x = "Time to First Failure", y = "Frequency") + 
  theme_minimal()
```
```{r}
head(result)
positive <- result$time_to_first_failure[result$time_to_first_failure > 0]
fit_norm <- fitdist(positive, "norm")
fit_weibull <- fitdist(positive, "weibull", method = "mle", lower = c(0, 0))
fit_gamma <- fitdist(positive, "gamma", method = "mle", lower = c(0, 0), start = list(scale = 1, shape = 1))
fit_lnorm <- fitdist(positive, "lnorm")

fit_norm
fit_weibull
fit_gamma
fit_lnorm
```
```{r}
ggplot(result, aes(x = time_to_first_failure)) +
  # Histogram
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "black", fill = "blue", position = "identity") +
  
  # Normal Distribution
  stat_function(
    fun = dnorm, 
    args = list(mean = fit_norm$estimate["mean"], sd = fit_norm$estimate["sd"]), 
    aes(color = "Normal"), size = 1
  ) +
  
  # Weibull Distribution
  stat_function(
    fun = dweibull, 
    args = list(shape = fit_weibull$estimate["shape"], scale = fit_weibull$estimate["scale"]), 
    aes(color = "Weibull"), size = 1
  ) +
  
  # Gamma Distribution
  stat_function(
    fun = dgamma, 
    args = list(shape = fit_gamma$estimate["shape"], rate = fit_gamma$estimate["rate"]), 
    aes(color = "Gamma"), size = 1
  ) +
  
  # Log-Normal Distribution
  stat_function(
    fun = dlnorm, 
    args = list(meanlog = fit_lnorm$estimate["meanlog"], sdlog = fit_lnorm$estimate["sdlog"]), 
    aes(color = "Log-Normal"), size = 1
  ) +
  
  # Labels and Theme
  labs(
    title = "Distribution of Time to First Failure",
    x = "Time to First Failure",
    y = "Density",
    color = "Distributions"  # Legend title
  ) +
  theme_minimal()
```
Visually speaking it looks as if this is a combination of several different distributions. What might be more useful is to do a segmented analysis so that we can fit distributions to parts of the network that might exhibit similar behaviour. The best way to do this would be to split the data based on material, and then further split it based on it's installation date (broad groups of installatio dates)


```{r}
# Generate observed densities 
observed <- hist(positive, breaks = 30, plot = F)
observed_x <- observed$mids
observed_y <- observed$density

#Compare predicted densities for each distribution 
predicted_norm <- dnorm(observed_x, mean = fit_norm$estimate["mean"], sd = fit_norm$estimate["sd"])
predicted_weibull <- dweibull(observed_x, shape = fit_weibull$estimate["shape"], scale = fit_weibull$estimate["scale"])
predicted_gamma <- dgamma(observed_x, shape = fit_gamma$estimate["scale"], rate = fit_gamma$estimate["shape"])
predicted_lnorm <- dlnorm(observed_x, meanlog = fit_lnorm$estimate["meanlog"], sdlog = fit_lnorm$estimate["sdlog"])

rmse <- function(observed, predicted) {
  sqrt(mean((observed-predicted)^2))
}

rmse_norm <- rmse(observed_y, predicted_norm)
rmse_weibull <- rmse(observed_y, predicted_weibull)
rmse_gamma <- rmse(observed_y, predicted_gamma)
rmse_lnorm <- rmse(observed_y, predicted_lnorm)



performance_metrics <- data.frame(
  Distribution = c("normal", "weibull", "gamma", "log_normal"), 
  RMSE = c(rmse_norm, rmse_weibull, rmse_gamma, rmse_lnorm)
)


performance_metrics
```

Doing a segemented time to first failure analysis, filtering it out based on age, diameter, material and installation date


```{r}
result$MATERIAL <- factor(result$MATERIAL)

MATERIALS <- split(result, result$MATERIAL)


AC.analysis <- MATERIALS$AC
ALK.analysis <- MATERIALS$ALK
CI.analysis <- MATERIALS$CI
CLS.analysis <- MATERIALS$CLS
CONC.analysis <- MATERIALS$CONC
CU.analysis <- MATERIALS$CU
DI.analysis <- MATERIALS$DI
MS.analysis <- MATERIALS$MS
PE.analysis <- MATERIALS$PE
PP.analysis <- MATERIALS$PP
PVCU.analysis <- MATERIALS$`PVC-U`




```




Defining a new function that produces 9 different plots. Each plot will have a sample distrubtion and several fitted distribution. 

```{r}

dist.fitter <- function(data){
  
  positive <- data$time_to_first_failure[data$time_to_first_failure > 0]
  fit_norm <- fitdist(positive, "norm")
  fit_weibull <- fitdist(positive, "weibull", method = "mle", lower = c(0, 0))
  fit_gamma <- fitdist(positive, "gamma", method = "mle", lower = c(0, 0), start = list(scale = 1, shape = 1))
  fit_lnorm <- fitdist(positive, "lnorm")
  
  
  
  
  plot <- ggplot(data, aes(x = positive)) +
  # Histogram
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "black", fill = "blue", position = "identity") +
  
  # Normal Distribution
  stat_function(
    fun = dnorm, 
    args = list(mean = fit_norm$estimate["mean"], sd = fit_norm$estimate["sd"]), 
    aes(color = "Normal"), size = 1
  ) +
  
  # Weibull Distribution
  stat_function(
    fun = dweibull, 
    args = list(shape = fit_weibull$estimate["shape"], scale = fit_weibull$estimate["scale"]), 
    aes(color = "Weibull"), size = 1
  ) +
  
  # Gamma Distribution
  stat_function(
    fun = dgamma, 
    args = list(shape = fit_gamma$estimate["shape"], rate = fit_gamma$estimate["rate"]), 
    aes(color = "Gamma"), size = 1
  ) +
  
  # Log-Normal Distribution
  stat_function(
    fun = dlnorm, 
    args = list(meanlog = fit_lnorm$estimate["meanlog"], sdlog = fit_lnorm$estimate["sdlog"]), 
    aes(color = "Log-Normal"), size = 1
  ) +
  
  # Labels and Theme
  labs(
    title = "Distribution of Time to First Failure",
    x = "Time to First Failure",
    y = "Density",
    color = "Distributions"  # Legend title
  ) +
  theme_minimal()
  
  
  observed <- hist(positive, breaks = 30, plot = F)
  observed_x <- observed$mids
  observed_y <- observed$density
  
  #Compare predicted densities for each distribution 
  predicted_norm <- dnorm(observed_x, mean = fit_norm$estimate["mean"], sd = fit_norm$estimate["sd"])
  predicted_weibull <- dweibull(observed_x, shape = fit_weibull$estimate["shape"], scale = fit_weibull$estimate["scale"])
  predicted_gamma <- dgamma(observed_x, shape = fit_gamma$estimate["scale"], rate = fit_gamma$estimate["shape"])
  predicted_lnorm <- dlnorm(observed_x, meanlog = fit_lnorm$estimate["meanlog"], sdlog = fit_lnorm$estimate["sdlog"])

  
  rmse_norm <- rmse(observed_y, predicted_norm)
  rmse_weibull <- rmse(observed_y, predicted_weibull)
  rmse_gamma <- rmse(observed_y, predicted_gamma)
  rmse_lnorm <- rmse(observed_y, predicted_lnorm)
  
  
  
  performance_metrics <- data.frame(
    Distribution = c("normal", "weibull", "gamma", "log_normal"), 
    RMSE = c(rmse_norm, rmse_weibull, rmse_gamma, rmse_lnorm)
  )
  
  print(plot)
  
  return(performance_metrics)
    
}





```

AC analysis
```{r}

#General 
dist.fitter(AC.analysis)


#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970
```



ALK analysis 
```{r}
#General
dist.fitter(ALK.analysis)
#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```


CI analysis
```{r}
#General
dist.fitter(CI.analysis)

#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```

CLS analysis
```{r}
#General

dist.fitter(CLS.analysis)
#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```


CONC analysis 
```{r}
dist.fitter(CONC.analysis)
#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```

CU analysis 
```{r}
dist.fitter(CU.analysis)
#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```

DI analysis 
```{r}

dist.fitter(DI.analysis)


#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```


MS analysis
```{r}

dist.fitter(MS.analysis)

#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```


PE analysis 
```{r}


dist.fitter(PE.analysis)

#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```



PP analysis
```{r} 

dist.fitter(PP.analysis)


#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```

PVC-U analysis
```{r}



dist.fitter(PVCU.analysis)
#Pipes pre 1930

#Pipes between 1930 and 1970 

#Pioes after 1970

```



## Correlation Analysis 


## Survival Analysis


## Key Customer Category Analysis


##Data Sources and Missing Data


## Confidence Sources Analysis